---
title: "Geospatial Project"
author: "Joan Costa Rodríguez, Francesco Zucca, Xènia Alemany Rovira, Michael Fehl"
date: today
format:
  pdf:
    toc: true
    number-sections: true
    geometry: margin=2.5cm
    fontsize: 12pt
    code-overflow: wrap
    syntax-highlighting: github
    
bibliography: "references.bib"
execute:
  echo: true
  warning: false
---

```{r libraries, results='hide'}
rm(list=ls())
library(readxl)        # read excel files
library(stringr)       # string operations
library(terra)         # raster operations
library(exactextractr) # exact_extract() for distances
library(gdistance)     # fast distance() formula for geo data
library(sf)            # vector operations
library(spData)        # geo data
library(tidyverse)     # NOTE: tidyr::extract() masks terra::extract()
library(ggrepel)       # add labels to ggplot
library(tidyterra)     # visualize geospatial using tidy syntax
library(patchwork)     # patch tg ggplot objects
library(did2s)
library(BMisc)
library(DRDID)
library(nprobust)   # Local polynomial regression for dose-response curve
library(cobalt)     # Covariate balance
library(stringi)
```

# Abstract

Does the severity of earthquakes determine how much additional remittance income a municipality receives after the disaster happened? Unlike existing macro-level studies that rely on state data, this paper exploits more granular municipal level data to evaluate the impact of earthquake severity on post-disaster remittances. At this local level spatial spillovers between neighboring municipalities may violate SUTVA, introducing attenuation bias and artificially dampening the observed regional differences.

To investigate this, we construct a municipality-quarter panel of Mexican remittance inflows from Banco de México (2013 - 2025) [@banxico2025], merged with Peak Ground Acceleration (PGA) data from the USGS Earthquake Hazard Program [@usgs2026]. Using this dataset, we estimate a Two-Way Fixed Effects (TWFE) Difference-in-Differences model - **actually we probably need to use Callaway Sant' Anna :(** -, where treatment intensity is defined as the average PGA experienced by each municipality.

# Introduction

In 2024 alone, Mexico received more that 67 billion current US dollars in remittances, making it the second country in the world by nominal remittance inflows, behind India [@worldbank2026]. When a shock hits the family in the municipality of origin, remittances act as an insurance mechanism, increasing flows for the following months [@lucas_stark1985, @bettin_jallow_zazzaro2023]. Mexico is one of the most seismically active countries in the world because of its location over three of the major tectonic plates [@rhea2011]. Particularly, on 7 September 2017, a magnitude 8.2 earthquake struck the coast of Chiapas, followed after around two weeks by a magnitude 7.1 event near Puebla, the latter one causing the collapse of 44 buildings and severely damaging another 600 [@singh2018].

[@mohapatra2012] use cross-country data and find that remittances increase after natural disasters, especially where there is a large stock of migrant population. [@bettin2025] use monthly bilateral remittance data from Italy to 81 countries and estimate an immediate 2% increase following these type of events, reverting back to the normal flow after around one year. While precise, these papers aggregate at the country level and rely on a binary treatment definition. Our approach, in contrast, utilizes more granular data and a continuous measure of treatment. This allows us to test two competing outcomes: either we will identify a stronger, localized response effect in the hardest-hit municipalities, or we will find that network effects completely absorb these granular differences.

# Data

## Remittance Data

The outcome of our analysis is quarterly remittance inflows from Banco de México SIE table CE166 - *Ingresos por Remesas, distribución por municipio* -, covering all Mexican municipalities from Q1 2013 to Q4 2025.

```{r remittance data}
# Read the excel skipping the metadata rows
remit_raw <- read_excel(
  "data/remittances/remittances.xlsx",
  skip = 8,
  col_types = "text"
)
```

We convert the dataset to long format and convert the date column in date format, adding columns for years and quarters to improve readability.

```{r remittance data long}
# Convert to long format
remit_long <- remit_raw %>%
  select(-`Information type`) %>% # Drop levels columns
  pivot_longer(
    cols      = -c(Title, Serie),
    names_to  = "quarter_label",
    values_to = "remittances_musd"
  ) %>%
  mutate(remittances_musd = as.numeric(remittances_musd))

# Convert quarter column to date format
remit_long <- remit_long %>%
  mutate(
    date    = as.Date(as.numeric(quarter_label), origin = "1899-12-30"),
    year    = as.integer(format(date, "%Y")),
    quarter = ceiling(as.integer(format(date, "%m")) / 3)
  ) %>%
  select(-quarter_label) # drop the old quarter column
```

Being interested only in municipality level data, we drop the State aggregate rows. Moreover, to improve readability once more, we add the corresponding state column to the left of the municipality one.

```{r remittance data cleaning}
# Tidy up municipality names
remit_long <- remit_long %>%
  # keep only remittances rows (drop state-level aggregates)
  filter(str_detect(Title, "Distribution by municipality")) %>%
  mutate(
    state        = str_trim(str_extract(Title, "(?<=, )[^,]+(?=, [^,]+$)")),
    municipality = str_trim(str_extract(Title, "[^,]+$"))
  ) %>%
  select(-Title)  # drop the original Title column

# Drop unused columns and rearrange columns
remit_long <- remit_long %>%
  select(-date, -Serie) %>%
  select(state, municipality, year, quarter, remittances_musd)
```

## Earthquake Data

We source ShakeMap raster data from the [USGS Earthquake Hazard Program](https://www.usgs.gov/programs/earthquake-hazards/earthquakes) for 10 significant seismic events (M6.7-M8.2) affecting Mexican territory between 2014 and 2022. For each event, we read the `HDF5` ShakeMap file, construct a georeferenced PGA raster, and compute municipality-level mean PGA via zonal statistics over INEGI municipal boundaries. We prefer PGA over earthquake magnitude, as PGA better captures actual ground motion experienced at each location, accounting for depth, distance, and local geology, making it comparable across events.

Our sample includes all events of Mw ≥ 6.7 affecting Mexican territory between 2014 and 2022 for which USGS ShakeMap raster data are available.

| Event | Date | Magnitude | Location |
|----|----|----|----|
| 2014 Coyuquilla Norte | April 18, 2014 | M7.2 | Guerrero |
| 2014 Puerto Madero | July 7, 2014 | M6.9 | Chiapas-Guatemala border |
| 2017 Chiapas | September 7, 2017 | M8.2 | Off coast of Chiapas |
| 2017 Matzaco (Puebla) | September 19, 2017 | M7.1 | Puebla/Morelos |
| 2018 Pinotepa | February 16, 2018 | M7.2 | Oaxaca |
| 2019 Puerto Madero | February 1, 2019 | M6.6-6.7 | Chiapas |
| 2020 Santa María Xadani | June 23, 2020 | M7.4 | Oaxaca |
| 2021 Acapulco | September 7, 2021 | M7.0 | Guerrero |
| 2022 Aguililla | September 19, 2022 | M7.6 | Michoacán |
| 2022 Aguililla (aftershock) | September 22, 2022 | M6.8 | Michoacán |

: \>6.5Mw Seismic events in Mexico, 2014-2022 {#tbl-earthquakes}

```{r, eval=FALSE}
# run script thru all earthquake data, manually updating script

```

```{r}
# check output data
df <- read.csv("output/2017_M8.2_chiapas/municipality_mean_intensity_pga.csv")
colnames(df)
# head(df)

# compare to remittance data to merge
colnames(remit_long)
# head(remit_long)
```

## Merging Datasets

We merge the remittance panel with the PGA exposure data on state x municipality x year x quarter. A direct string match yields 82% of municipalities matched. The remaining mismatches stem from systematic naming differences between Banxico remittance data and the INEGI earthquake shapefile. After manual renaming, 2,427 of 2,431 municipalities are matched, with 4 remaining unmatched due to shapefile version discrepancies. For municipality-quarters with no earthquake, PGA is set to zero.

### Data Cleaning: Name Mismatches

First, we try just a direct string join between the two datasets based on corresponding state & municipality name matches.

```{r}
## test matching muni names

# load one PGA file
pga <- read.csv("output/2017_M8.2_chiapas/municipality_mean_intensity_pga.csv") %>%
  select(adm1_name, adm2_name, adm2_pcode, mean_pga)

# attempt a direct join
test_join <- remit_long %>%
  distinct(state, municipality) %>%
  left_join(pga, by = c("state" = "adm1_name", "municipality" = "adm2_name"))

# How many matched vs failed?
cat("Total municipalities in remittances:", nrow(test_join), "\n")
cat("Matched:", sum(!is.na(test_join$adm2_pcode)), "\n")
cat("Unmatched:", sum(is.na(test_join$adm2_pcode)), "\n")  # 450 mismatches
```

We return a mismatch of 450 total municipalities, meaning \~20% of our data is not aligned. We suspect this has to do with a divergence in how the names of certain Mexican states and municipalities are recorded by different administrations.

```{r, eval=FALSE, echo=FALSE}
# check and see what munis are missing from merge
unmatched <- test_join %>% filter(is.na(adm2_pcode))

unmatched %>% 
  select(state, municipality) %>% 
  arrange(state) %>% 
  print(n = 25)

# different names: accents, etc
pga %>% 
  filter(grepl("Coahuila", adm1_name)) %>% 
  distinct(adm1_name)

pga %>% filter(grepl("México|Ciudad|Distrito", 
                     adm1_name, 
                     ignore.case = TRUE)) %>% 
  distinct(adm1_name)

# how many are "Unidentified"
unmatched %>% 
  filter(municipality == "Unidentified") %>% 
  nrow()

# how many are real municipalities
unmatched %>% 
  filter(municipality != "Unidentified") %>% 
  nrow()

# what munis are not aligning, sorted by state
unmatched %>% 
  filter(municipality != "Unidentified") %>%
  count(state, sort = TRUE) %>% 
  print(n = 32)

# state name mismatch: Veracruz & michoacan
pga %>% 
  filter(grepl("Veracruz|Michoacán", adm1_name)) %>% 
  distinct(adm1_name)

# state name mismatch: Queretaro
pga %>% filter(grepl("Quer", adm1_name)) %>% 
  distinct(adm1_name)


# check Nuevo Leon spellings
pga %>% filter(adm1_name == "Nuevo León", 
               grepl("General|Doctor|Carmen", adm2_name)) %>% 
  select(adm1_name, adm2_name)

# check the others edge cases
pga %>% filter(adm1_name %in% c("Guanajuato","Jalisco","Morelos","México",
                                "Oaxaca","Tlaxcala","Chiapas"),
               grepl("Silao|Tlaquepaque|Tlaltizapán|Acambay|Teoloyucán|Mixtepec|
                     Texmelucan|Tezoatlán|Zapotitlán|Altzayanca|Nicolás|Dolores", 
                     adm2_name, ignore.case = TRUE)) %>%
  select(adm1_name, adm2_name)


pga %>% filter(grepl("Teoloyucán|Texmelucan|Zapotitlán del|Altzayanca", 
                      adm2_name, ignore.case = TRUE)) %>%
  select(adm1_name, adm2_name)
```

```{r, echo=FALSE}
# fix spellings in remittance municipalities
remit_long <- remit_long %>%
  mutate(municipality = case_when(
    # Nuevo Leon abbreviations
    state == "Nuevo León" & municipality == "Dr. Arroyo"      ~ "Doctor Arroyo",
    state == "Nuevo León" & municipality == "Dr. Coss"        ~ "Doctor Coss",
    state == "Nuevo León" & municipality == "Dr. González"    ~ "Doctor González",
    state == "Nuevo León" & municipality == "Carmen"          ~ "El Carmen",
    state == "Nuevo León" & municipality == "Gral. Bravo"     ~ "General Bravo",
    state == "Nuevo León" & municipality == "Gral. Escobedo"  ~ "General Escobedo",
    state == "Nuevo León" & municipality == "Gral. Terán"     ~ "General Terán",
    state == "Nuevo León" & municipality == "Gral. Treviño"   ~ "General Treviño",
    state == "Nuevo León" & municipality == "Gral. Zaragoza"  ~ "General Zaragoza",
    state == "Nuevo León" & municipality == "Gral. Zuazua"    ~ "General Zuazua",
    # Others
    state == "México"     & municipality == "Acambay"         ~ "Acambay de Ruíz Castañeda",
    state == "Oaxaca"     & municipality == "Tezoatlán de Segura y Luna" ~ "Heroica Villa Tezoatlán de Segura y Luna",
    state == "Jalisco"    & municipality == "Tlaquepaque"     ~ "San Pedro Tlaquepaque",
    state == "Guanajuato" & municipality == "Silao"           ~ "Silao de la Victoria",
    state == "Morelos"    & municipality == "Tlaltizapán"     ~ "Tlaltizapán de Zapata",
    state == "Chiapas"    & municipality == "Nicolás Ruiz"    ~ "Nicolás Ruíz",
    state == "Guanajuato" & municipality == "Dolores Hidalgo Cuna de la Independencia Nal." ~ "Dolores Hidalgo Cuna de la Independencia Nacional",
    state == "Oaxaca" & municipality == "San Juan Mixtepec - Distr. 08 -"  ~ "San Juan Mixtepec -Dto. 08 -",
    state == "Oaxaca" & municipality == "San Juan Mixtepec - Distr. 26 -"  ~ "San Juan Mixtepec -Dto. 26 -",
    state == "Oaxaca" & municipality == "San Pedro Mixtepec - Distr. 22 -" ~ "San Pedro Mixtepec -Dto. 22 -",
    state == "Oaxaca" & municipality == "San Pedro Mixtepec - Distr. 26 -" ~ "San Pedro Mixtepec -Dto. 26 -",
    TRUE ~ municipality
  ))

# Drop unidentified rows
remit_long <- remit_long %>% filter(municipality != "Unidentified")


# Fix state names in PGA data to match remittance data
pga <- pga %>%
  mutate(adm1_name = case_when(
    adm1_name == "Coahuila de Zaragoza"            ~ "Coahuila",
    adm1_name == "Distrito Federal"                ~ "Ciudad de México",
    adm1_name == "Veracruz de Ignacio de la Llave" ~ "Veracruz",
    adm1_name == "Michoacán de Ocampo"             ~ "Michoacán",
    adm1_name == "Querétaro de Arteaga"            ~ "Querétaro",
    TRUE                                            ~ adm1_name
  ))

# merge with remittances
test_join <- remit_long %>%
  distinct(state, municipality) %>%
  left_join(pga, by = c("state" = "adm1_name", "municipality" = "adm2_name"))

# How many matched vs failed?
cat("Matched:", sum(!is.na(test_join$adm2_pcode)), "\n")
cat("Unmatched:", sum(is.na(test_join$adm2_pcode)), "\n") # 4 mismatches (can leave NA)
```

After identifying divergent spellings and correcting them in order to match properly, we return a final merged dataset consisting of only four mismatched municipalities, meaning only a 0.16% mismatch in names.

### Data Pipeline

Finally, we implement this data cleaning pipeline for all ten of our recorded earthquakes, stacking the resulting municipality-level PGA estimates into a single exposure dataset and merging it with the remittance panel on state, municipality, year, and quarter. For quarters in which no earthquake occurred, the PGA value is set to zero. Where two earthquakes share the same quarter, we retain the maximum PGA value across events, as the two 2017 events and two 2022 events affect geographically distinct regions of Mexico with negligible overlap at damage-relevant thresholds.

```{r}
# step 1: store all earthquake folders in an object
events <- c(
  "2014_M6.9_puerto_madero",
  "2014_M7.2_coyuquilla_norte",
  "2017_M7.1_matzaco",
  "2017_M8.2_chiapas",
  "2018_M7.2_pinotepa",
  "2019_M6.7_puerto_madero",
  "2020_M7.4_Santa_Maria_Xadani",
  "2021_M7.0_acapulco",
  "2022_M6.8_aguililla",
  "2022_M7.6_aguililla"
)

# step 2: read PGA csv files from each folder
pga_all <- map_dfr(events, function(ev) { 
  read.csv(paste0("output/", ev, "/municipality_mean_intensity_pga.csv")) %>%
    mutate(event = ev) %>%
    select(adm1_name, adm2_name, mean_pga, event)
})

# apply state name fixes
pga_all <- pga_all %>%
  mutate(adm1_name = case_when(
    adm1_name == "Coahuila de Zaragoza"            ~ "Coahuila",
    adm1_name == "Distrito Federal"                ~ "Ciudad de México",
    adm1_name == "Veracruz de Ignacio de la Llave" ~ "Veracruz",
    adm1_name == "Michoacán de Ocampo"             ~ "Michoacán",
    adm1_name == "Querétaro de Arteaga"            ~ "Querétaro",
    TRUE                                            ~ adm1_name
  ))

# step 3: store event date/quarter info, matching remittance data structure
event_dates <- tribble(
  ~event,                        ~year, ~quarter,
  "2014_M6.9_puerto_madero",     2014,  3,
  "2014_M7.2_coyuquilla_norte",  2014,  2,
  "2017_M7.1_matzaco",           2017,  3,
  "2017_M8.2_chiapas",           2017,  3,
  "2018_M7.2_pinotepa",          2018,  1,
  "2019_M6.7_puerto_madero",     2019,  1,
  "2020_M7.4_Santa_Maria_Xadani",2020,  2,
  "2021_M7.0_acapulco",          2021,  3,
  "2022_M6.8_aguililla",         2022,  3,
  "2022_M7.6_aguililla",         2022,  3
)

# step 4: merge in YQ info into PGA data
pga_all <- pga_all %>% 
  left_join(event_dates, by = "event")

# step 5: merge PGA with remittance panel
panel <- remit_long %>%
  left_join(
    pga_all,
    by = c("state" = "adm1_name", 
           "municipality" = "adm2_name",
           "year", "quarter")
  ) %>%
  mutate(mean_pga = replace_na(mean_pga, 0))
```

```{r}
# sanity check
head(panel, n=5) # should be returning us ~126,412 rows (# of remit. rows)

## issue: both 2017 earthquakes in Q3, and both 2022 earthquakes in Q3
panel %>% 
  slice(19:20)

# gives us duplicate rows with same values
panel %>% 
  filter(year == 2017, quarter == 3) %>%
  count(state, municipality, sort = TRUE) %>%
  filter(n > 1)
```

Note that after the final merging of remittance data to the PGA data, we find after careful investigation duplicate rows for each municipality where there is two different earthquakes within the same quarter (2017Q3 and 2022Q3).

```{r}
# solution: filter only the most relevant (geographically) earthquake for a muni
panel_check <- panel %>%
  filter(year == 2017, quarter == 3) %>%
  group_by(state, municipality) %>%
  summarise(
    pga_M8.2 = mean_pga[event == "2017_M8.2_chiapas"],
    pga_M7.1 = mean_pga[event == "2017_M7.1_matzaco"],
    .groups = "drop"
  ) %>%
  filter(pga_M8.2 > 5 & pga_M7.1 > 5)  
# municipalities hit by both earthquakes (PGA > 5%g)

nrow(panel_check)

# applied to our data set
panel_filtered <- panel %>%
  group_by(state, municipality, year, quarter, remittances_musd) %>%
  slice_max(mean_pga, n = 1, with_ties = FALSE) %>%
  ungroup() %>% 
  mutate(remittances_musd=remittances_musd*1000)



# use max() to select most relevant earthquake

nrow(panel_filtered)

# store as CSV file
write.csv(panel_filtered, "data/panel_remittances_pga.csv", row.names = FALSE)
```

# Spillover Effects

## Migration Networks

To get a measure of potential spillover effects, we can use a manually constructed weighting matrix of mexican migration from a given Mexican municipality to each US state (computed via MCAS migration data from 2010-2024). We can then multiply this matrix by its transpose to get a municipality x municipality matrix to try and understand the connectedness of certain Mexican municipalities based on their shares US migration destinations. Theoretically, this means that certain Mexican municipalities that are not close in geographic distance can still propagate shocks through these migration networks.

```{r import weighting matrix}
M <- read_excel("data/AVG_WEIGHTING_MATRIX.xlsx") 
muni_ids <- M[[2]]   # store IDs
state_ids <- M[[1]]  
muni_labels <- paste(state_ids, muni_ids, sep = "_") 
# generate unique label for each muni

M <- as.matrix(M[c(-1, -2)]) # drop names, convert to matrix
dim(M) # (2638 x 51) <==> (MX muni x US states) (District of Columbia gives us +1)

# transpose matrix
M_prime <- t(M)

# MM' gives our (muni x muni) network
W_network <- M %*% M_prime
dim(W_network)

# place muni names back in
rownames(W_network) <- muni_labels
colnames(W_network) <- muni_labels

# municipality is not its own neighbor
diag(W_network) <- 0

# row normalize: add to one
W_network <- W_network/rowSums(W_network)

# sanity check
as.data.frame(W_network) %>% 
  slice(1:4) %>% 
  select(1:2)

# largest values in a row = strongest network neighbors
# sort(W_network["Michoacan_Coeneo",], decreasing = TRUE)[1:5]

# Persist W_network so the CS section can use it after rm(list=ls())
saveRDS(W_network, "data/W_network.rds")
```


```{r clean up environment, results='hide', echo=FALSE}
rm(M, M_prime, muni_ids, state_ids)
```

This spillover matrix serves to approximately measure the "connectedness" between every pair of municipalities in Mexico, based on their historical migration patterns to the US. The idea is that if an earthquake hits Asientos in Aguascalientes, the migration community in (say) Texas that comes from this municipality will send more money home. But if that same community in Texas also has a large network of migrants from Coeneo in Michoacan, they might discuss the earthquake; consequently, migrants from Coeneo may send additional remittances back home as feel a heightened sense of their own family's vulnerability.

We suspect this effect, if present, would be stronger among municipalities sharing the same US migration destinations, as information travels through co-ethnic social networks rather than purely along geographic borders. We test this by estimating whether 'never-treated' municipalities with stronger network overlap with earthquake-affected areas experienced higher remittance inflows in the post-event period, conditional on municipality and time fixed effects.

## Geographic Distance

Furthermore, geographic proximity may independently drive spillover remittances. Even absent shared migration networks, US-based migrants aware of a nearby earthquake may send precautionary remittances to family in adjacent municipalities, which would reflect a generalized regional concern, rather than direct social-network linkages. To test this, we construct an inverse-distance weight matrix $W^{\text{geo}}$ using municipal centroids and compute a geographic PGA exposure variable for each never-treated municipality.

To do so, we can simply find the centroid of each municipality using the `mex_admin2.shp` file, as it contains each municipality's associated geometry. Note that we define the WGS to standard `EPSG:4326` to be consistent. Afterwards, we just calculate all the pairwise distances between each centroid to generate a distance matrix (in meters).

```{r geographic matrix}
# municipal boundaries data
mex_muni <- st_read("data/earthquakes/mex_admin2/mex_admin2.shp", 
                    quiet = TRUE) %>%
  st_transform("EPSG:4326") %>%
  mutate(adm1_name = case_when(
    adm1_name == "Coahuila de Zaragoza"            ~ "Coahuila",
    adm1_name == "Distrito Federal"                ~ "Ciudad de México",
    adm1_name == "Veracruz de Ignacio de la Llave" ~ "Veracruz",
    adm1_name == "Michoacán de Ocampo"             ~ "Michoacán",
    adm1_name == "Querétaro de Arteaga"            ~ "Querétaro",
    TRUE                                            ~ adm1_name
  )) %>%
  mutate(muni_id = paste(adm1_name, adm2_name, sep = "_"))

muni_centroids <- st_centroid(mex_muni)         # centroid of each muni
dist_mat       <- st_distance(muni_centroids)   # compute all pairwise distances
```

Now we can standardize this matrix to measure 'intensity' by rescaling all our values to become $\in [0,1]$. Values closer to 1 are closer in geographic distance, relative to all other municipality distances. 

```{r}
# diagonal = Inf so that 1/Inf = 0 (municipality is not its own neighbor)
diag(dist_mat) <- Inf
W_geo          <- 1 / dist_mat
diag(W_geo)    <- 0

# row-normalize: each row sums to 1
W_geo <- W_geo / rowSums(W_geo)

# name rows/cols with state_muni so they match muni_id in the panel
rownames(W_geo) <- mex_muni$muni_id
colnames(W_geo) <- mex_muni$muni_id

# store as network
saveRDS(W_geo, "data/W_geo.rds")
```


# Regression

## Parallel Trends Assumption

Before estimating the model, we verify the parallel trends assumption visually by comparing average remittance trajectories between ever-treated and never-treated municipalities in the pre-treatment period.

```{r, echo=FALSE}
rm(list=ls())
panel <- read_csv("data/panel_remittances_pga.csv")

# PARALLEL TRENDS ASSUMPTION

# define 'ever-treated' municipalities
ever_treated <- panel %>%
  group_by(state, municipality) %>%
  summarise(ever_treated = max(mean_pga, na.rm = TRUE) > 2, 
            .groups = "drop")

# plot
panel %>%
  left_join(ever_treated, by = c("state", "municipality")) %>%
  mutate(treated = ifelse(ever_treated, "Ever Treated", "Never Treated")) %>%
  group_by(year, quarter, treated) %>%
  summarise(avg_remit = mean(remittances_musd, na.rm = TRUE), .groups = "drop") %>%
  mutate(time = year + (quarter - 1) / 4) %>%
  ggplot(aes(x = time, y = avg_remit, color = treated)) +
  geom_line() +
  geom_vline(xintercept = 2017.5, color = "red", linetype = "dashed") +
  labs(title = "Remittances: ever-treated vs never-treated municipalities",
       x = "Year", y = "Avg remittances (M USD)")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

# notes: 
# - parallel trends seems to hold
# - but no obvious divergence; makes sense as we are looking at the whole picture; at the aggregate level
#     - i.e. this averages across all treated municipalities regardless of PGA intensity
# - would expect DiD regression to be much more precise; PGA will be used as a continuous variable

```

The pre-treatment trends appear broadly parallel, providing visual support for the parallel trends assumption, though we cannot rule out pre-existing differences between treated and control municipalities.

## Panel Preparation

```{r panel-prep}
library(tidyverse)
library(fixest)
library(did)

# 1. Build panel
panel <- panel %>%          # use the deduplicated panel (max PGA)
  rename(event_quarter = event) %>%
  mutate(
    muni_id    = paste(state, municipality, sep = "_"),
    time_index = (year - min(year)) * 4 + quarter,
    asinhremit  = asinh(remittances_musd)
  )

# 2. PGA partitions
partitions <- list(
  "4-10"  = c(4, 10),
  "10-20" = c(10, 20),
  "20+"   = c(20, Inf)
)

PGA_THRESHOLD <- 4   # municipalities below this are controls

# 3. Municipality-level metadata 
# uses slice_max on mean_pga (strongest shock) 
muni_meta <- panel %>%
  filter(!is.na(event_quarter), mean_pga > PGA_THRESHOLD) %>%
  group_by(muni_id) %>%
  slice_max(mean_pga, n = 1, with_ties = FALSE) %>%
  transmute(
    muni_id,
    event = event_quarter,
    dose  = mean_pga,
    G     = time_index
  ) %>%
  ungroup()

# 4. Merge back 
panel <- panel %>%
  left_join(muni_meta, by = "muni_id")
# event=NA after join gives us a clean 'never-treated' control

# 5. Sanity check
earthquake_cohorts <- muni_meta %>%
  distinct(event, G) %>%
  arrange(G)

cat("Earthquake cohorts:\n")
print(earthquake_cohorts)

cat("\nTreated municipalities per cohort:\n")
muni_meta %>% count(event, G, name = "n_treated") %>% arrange(G) %>% print()

cat("\nNever-treated controls:", 
    panel %>% filter(is.na(event)) %>% distinct(muni_id) %>% nrow(), "\n")
```


## Callaway and Sant'Anna

We prepare the panel and partition the sample into three categories. For each partition, we compute the Callaway and Sant'Anna estimand and examine whether the estimated effect differs according to the level of PGA received by the municipality.

```{r}

panel_cs <- panel %>%
  mutate(
    G_cs     = ifelse(is.na(event), 0L, as.integer(G)),
    muni_num = as.integer(factor(muni_id))
  )
 
# Override partitions: merge 20-35 and 35+ into a single 20+ bin
partitions <- list(
  "4-10"  = c(4, 10),
  "10-20" = c(10, 20),
  "20+"   = c(20, Inf)
)
```

We select a pre-window of −10 periods to test whether the parallel trends assumption holds, and a post-window of 10 periods to examine whether the effects on remittances are not immediate but instead materialize with a lag that may not be apparent at first glance.

```{r}

run_cs_partition <- function(data, pga_lo, pga_hi, partition_label,
                              pre_window  = -10,
                              post_window =  10) {
 
  treated_in_partition <- data %>%
    filter(!is.na(event), dose > pga_lo, dose <= pga_hi) %>%
    distinct(muni_id)
 
  if (nrow(treated_in_partition) == 0) {
    cat("\nPartition", partition_label, "→ SKIPPED: no treated municipalities\n")
    return(NULL)
  }
 
  sample_df <- data %>%
    filter(
      muni_id %in% treated_in_partition$muni_id |
      G_cs == 0
    ) %>%
    mutate(
      G_cs = ifelse(muni_id %in% treated_in_partition$muni_id, G_cs, 0L)
    )
 
  n_treated  <- n_distinct(sample_df$muni_id[sample_df$G_cs > 0])
  n_controls <- n_distinct(sample_df$muni_id[sample_df$G_cs == 0])
  n_cohorts  <- n_distinct(sample_df$G_cs[sample_df$G_cs > 0])
 
  if (n_treated < 2) {
    cat("  → SKIPPED: fewer than 2 treated municipalities\n")
    return(NULL)
  }
 
  cs_out <- tryCatch(
    att_gt(
      yname         = "asinhremit",
      tname         = "time_index",
      idname        = "muni_num",
      gname         = "G_cs",
      data          = sample_df,
      control_group = "nevertreated",
      est_method    = "dr",
      clustervars   = "muni_num",
      print_details = FALSE
    ),
    error = function(e) {
      cat("  → ERROR in att_gt:", conditionMessage(e), "\n")
      NULL
    }
  )
 
  if (is.null(cs_out)) return(NULL)
 
  es <- tryCatch(
    aggte(cs_out,
          type  = "dynamic",
          min_e = pre_window,
          max_e = post_window,
          na.rm = TRUE),
    error = function(e) {
      cat("  → ERROR in aggte(dynamic):", conditionMessage(e), "\n"); NULL
    }
  )
 
  agg <- NULL
  if (!is.null(es)) {
    post_idx <- which(es$egt >= 0 & es$egt <= post_window)
    if (length(post_idx) > 0) {
      post_atts <- es$att.egt[post_idx]
      post_att  <- mean(post_atts, na.rm = TRUE)
      post_se   <- sqrt(mean(es$se.egt[post_idx]^2, na.rm = TRUE) / length(post_idx))
      agg <- list(overall.att = post_att, overall.se = post_se)
    }
  }
 
  list(
    partition  = partition_label,
    n_treated  = n_treated,
    n_controls = n_controls,
    n_cohorts  = n_cohorts,
    cs_out     = cs_out,
    agg        = agg,
    es         = es
  )
}
```

We run the Callaway & Sant'Anna estimator for each PGA partition. The "map()" call iterates over the three partitions defined earlier, invoking run_cs_partition() for each one with the corresponding PGA bounds and the pre/post windows specified above

```{r}
cs_results <- map(names(partitions), function(part_name) {
  bounds <- partitions[[part_name]]
  run_cs_partition(
    data           = panel_cs,
    pga_lo         = bounds[1],
    pga_hi         = bounds[2],
    partition_label = part_name,
    pre_window     = -10,
    post_window    =  10
  )
}) %>%
  setNames(names(partitions)) %>%
  compact()
```

We then collect the aggregate ATT estimates into a single tidy dataframe. For each partition, we extract the overall ATT, its standard error, the implied 95% confidence interval, and a two-sided p-value derived from the normal approximation. Significance stars are appended following conventional thresholds (10%, 5%, 1%), and partitions are recast as an ordered factor to ensure correct sorting in downstream tables and plots.

```{r}
att_table <- map_dfr(cs_results, function(res) {
  if (is.null(res$agg)) return(NULL)
  tibble(
    partition  = res$partition,
    n_cohorts  = res$n_cohorts,
    n_treated  = res$n_treated,
    n_controls = res$n_controls,
    att        = res$agg$overall.att,
    se         = res$agg$overall.se,
    ci_lo      = res$agg$overall.att - 1.96 * res$agg$overall.se,
    ci_hi      = res$agg$overall.att + 1.96 * res$agg$overall.se,
    p_val      = 2 * pnorm(-abs(res$agg$overall.att / res$agg$overall.se))
  )
}) %>%
  mutate(
    sig       = case_when(
      p_val < 0.01 ~ "***",
      p_val < 0.05 ~ "**",
      p_val < 0.10 ~ "*",
      TRUE         ~ ""
    ),
    partition = factor(partition, levels = c("4-10", "10-20", "20+"))
  ) %>%
  arrange(partition)
 
print(att_table)
```

For a cleaner presentation, we can reshape the att_table:

```{r}
att_table %>%
  transmute(
    `PGA Partition`  = partition,
    `Cohorts`        = n_cohorts,
    `Treated Munis`  = n_treated,
    `Control Munis`  = n_controls,
    `ATT (0 to +4)`  = round(att,  4),
    `Std. Error`     = round(se,   4),
    `95% CI`         = paste0("[", round(ci_lo,4), ", ", round(ci_hi,4), "]"),
    `p-value`        = round(p_val, 3),
    ` `              = sig
  ) %>%
  knitr::kable(
    booktabs = TRUE,
    align    = c("l","c","c","c","r","r","c","r","l"),
    caption  = "CS ATT by PGA Partition | quarters 0 to +4 | doubly robust | never-treated controls"
  )
```

The results reveal a striking and economically meaningful pattern of heterogeneity across PGA partitions. For municipalities exposed to low- and moderate-intensity seismic shocks (PGA 4-10 and 10-20 %g), we estimate large, positive, and precisely estimated average treatment effects on remittances, ATTs of 0.513 and 0.496 log points, respectively, both significant at the 1% level. This finding is consistent with the consumption-smoothing motive for remittances: when a household in the origin municipality is struck by a moderate earthquake, migrants abroad respond by increasing transfers, effectively acting as informal insurers against idiosyncratic shocks and dampening the impact on consumption.

The picture changes heavyly for municipalities in the highest PGA partition (above 20 %g). Here, the estimated ATT turns negative (−0.506 log points) and remains statistically significant at the 5% level, also with a substantially larger standard error, reflecting the reduced sample of treated units in this category. This reversal is consistent with the migration-response mechanism documented by Uriel (2022), when a disaster is severe enough, affected households may themselves migrate to smooth consumption, thereby disrupting the pre-existing remittance relationship and potentially reversing the direction of transfers or reducing the stock of recipients in the origin municipality altogether.

Taken together, these results suggest that the remittance response to seismic shocks is non-monotonic in shock severity, with consumption smoothing via inflows dominating at moderate intensities and outmigration effects dominating at the upper tail of the shock distribution.

We can also complement the results with plots
```{r}
#| fig.width: 6
#| fig.height: 4
#| fig.align: "center"
#| out.width: "90%"

es_plots <- map(cs_results, function(res) {
  if (is.null(res$es)) return(NULL)

  ggdid(res$es) +
    labs(
      title    = paste("Event Study - PGA Partition:", res$partition),  # plain dash, safer for LaTeX
      subtitle = paste0("CS (2021) | never-treated controls | ",
                        res$n_treated, " treated, ",
                        res$n_controls, " control munis"),
      x        = "Quarters relative to earthquake",
      y        = "ATT (Inverse hyperbolic sine remittances)"
    ) +
    theme_bw(base_size = 11) +                        # base_size prevents tiny text in PDF
    theme(plot.title    = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5, size = 8),
          legend.position = "bottom")                 # avoids legend clipping on narrow pages
}) %>%
  compact()

walk(es_plots, print)
```
The event-study plots indicate that the parallel trends assumption is satisfied across all three partitions, with only  minor deviations observed in the lowest PGA group during the pre-treatment window, a reassuring diagnostic given the heterogeneity in seismic exposure across municipalities.

Beyond the pre-trends check, the plots reveal that the remittance response is not immediate but instead builds gradually over time. For municipalities in the low- and moderate-intensity partitions, the ATT follows a positive and increasing trajectory in the post-treatment periods, suggesting that migrant senders adjust their transfer behavior with a lag, consistent with the time it takes for information about the severity of the shock to reach the senders and for the corresponding financial response to materialize.

The pattern for the high-intensity partition tells a different story. While the point estimates turn negative in the first periods following the shock, in line with the aggregate ATT discussed above, the confidence intervals widen considerably as the horizon extends, reflecting a loss of precision rather than a reversal of the effect per se.

We also produce a combined dot-and-whisker plot that places the three aggregate ATT estimates side by side, allowing direct visual comparison across PGA partitions

```{r}
ggplot(att_table, aes(x = partition, y = att)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi),
                width = 0.15, color = "steelblue", linewidth = 0.8) +
  geom_point(color = "steelblue", size = 3) +
  geom_text(aes(label = sig, y = ci_hi + 0.01),
            size = 5, color = "steelblue") +
  labs(
    title    = "CS ATT by PGA Partition",
    subtitle = "Callaway & Sant'Anna (2021) | quarters 0 to +4 | doubly robust | never-treated controls",
    x        = "PGA Partition (%g)",
    y        = "ATT (Inverse hyperbolic sine remittances)"
  ) +
  theme_bw() +
  theme(plot.title    = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```


### Robustness check

The following specification replaces the never-treated municipalities with the lowest-dose treated group (PGA 4-10 %g) as the control group. The motivation is to restrict the comparison to municipalities that lie within the vicinity of an earthquake, and therefore share similar geographic and structural characteristics.

Using never-treated municipalities as controls risks confounding the earthquake effect with pre-existing differences between seismically active regions and the rest of the country, differences that could manifest as divergent trends in remittance inflows unrelated to the shock itself. By benchmarking higher-dose municipalities against those that experienced only a mild ground shaking, we hold constant the broad exposure to a seismic event and isolate the incremental effect of stronger shaking intensity on remittance inflows.

We follow a similar code structure as before but changing the status of the lower PGA group:

```{r}
panel_cs_ld <- panel %>%
  mutate(
    # Low-dose treated (4,10] → control group (G = 0)
    G_cs = case_when(
      is.na(event)                      ~ 0L,
      dose >  4 & dose <= 10            ~ 0L,   # ← low-dose = control
      TRUE                              ~ as.integer(G)
    ),
    muni_num = as.integer(factor(muni_id))
  )
 
# New partitions: only two treatment bins, both benchmarked against 4-10
partitions_ld <- list(
  "10-20" = c(10, 20),
  "20+"   = c(20, Inf)
)
 

panel_cs_ld %>%
  filter(!is.na(event)) %>%
  distinct(muni_id, dose) %>%
  mutate(group = case_when(
    dose >  4  & dose <= 10 ~ "4-10 (CONTROL)",
    dose > 10  & dose <= 20 ~ "10-20 (treated)",
    dose > 20               ~ "20+   (treated)",
    TRUE                    ~ "below 4 (excluded)"
  )) %>%
  count(group, sort = FALSE) %>%
  print()

run_cs_partition_ld <- function(data, pga_lo, pga_hi, partition_label,
                                 pre_window  = -10,
                                 post_window =  10) {
 
  # Treated municipalities in this partition
 
  treated_in_partition <- data %>%
    filter(!is.na(event), dose > pga_lo, dose <= pga_hi) %>%
    distinct(muni_id)
 
  if (nrow(treated_in_partition) == 0) {
    cat("\nPartition", partition_label, "→ SKIPPED: no treated municipalities\n")
    return(NULL)
  }
 
  # Controls = never-treated (no earthquake) + low-dose (4,10] — all have G_cs == 0
  # Drop municipalities treated in OTHER high-dose partitions (clean comparison)
  other_high_dose <- data %>%
    filter(!is.na(event), dose > 10,                       # high-dose treated
           !(muni_id %in% treated_in_partition$muni_id)) %>%  # but not in THIS partition
    distinct(muni_id)
 
  sample_df <- data %>%
    filter(
      muni_id %in% treated_in_partition$muni_id |   # this partition's treated
      G_cs == 0                                      # never-treated + low-dose controls
    ) %>%
    filter(!(muni_id %in% other_high_dose$muni_id)) %>%   # drop other partitions
    mutate(
      G_cs = ifelse(muni_id %in% treated_in_partition$muni_id, G_cs, 0L)
    )
 
  n_treated  <- n_distinct(sample_df$muni_id[sample_df$G_cs > 0])
  n_controls <- n_distinct(sample_df$muni_id[sample_df$G_cs == 0])
  n_cohorts  <- n_distinct(sample_df$G_cs[sample_df$G_cs > 0])
 

  if (n_treated < 2) {
    cat("  → SKIPPED: fewer than 2 treated municipalities\n")
    return(NULL)
  }

cs_out <- tryCatch(
    att_gt(
      yname         = "asinhremit",
      tname         = "time_index",
      idname        = "muni_num",
      gname         = "G_cs",
      data          = sample_df,
      control_group = "nevertreated",   # "nevertreated" now includes low-dose!
      est_method    = "dr",
      clustervars   = "muni_num",
      print_details = FALSE
    ),
    error = function(e) {
      cat("  → ERROR in att_gt:", conditionMessage(e), "\n"); NULL
    }
  )
 
  if (is.null(cs_out)) return(NULL)
 
  es <- tryCatch(
    aggte(cs_out,
          type  = "dynamic",
          min_e = pre_window,
          max_e = post_window,
          na.rm = TRUE),
    error = function(e) {
      cat("  → ERROR in aggte(dynamic):", conditionMessage(e), "\n"); NULL
    }
  )
 
  agg <- NULL
  if (!is.null(es)) {
    post_idx <- which(es$egt >= 0 & es$egt <= post_window)
    if (length(post_idx) > 0) {
      post_atts <- es$att.egt[post_idx]
      post_att  <- mean(post_atts, na.rm = TRUE)
      post_se   <- sqrt(mean(es$se.egt[post_idx]^2, na.rm = TRUE) / length(post_idx))
      agg <- list(overall.att = post_att, overall.se = post_se)
    }
  }
 
  list(
    partition  = partition_label,
    n_treated  = n_treated,
    n_controls = n_controls,
    n_cohorts  = n_cohorts,
    cs_out     = cs_out,
    agg        = agg,
    es         = es
  )
}
```

Run the function for every PGA partition:


```{r, warning=FALSE}
cs_results_ld <- map(names(partitions_ld), function(part_name) {
  bounds <- partitions_ld[[part_name]]
  run_cs_partition_ld(
    data            = panel_cs_ld,
    pga_lo          = bounds[1],
    pga_hi          = bounds[2],
    partition_label = part_name,
    pre_window      = -10,
    post_window     =  10
  )
}) %>%
  setNames(names(partitions_ld)) %>%
  compact()
 
cat("\n=== Partitions estimated successfully:",
    paste(names(cs_results_ld), collapse = ", "), "===\n")
```

As before, we construct a summary table with the results of the CS estimand:

```{r}
att_table_ld <- map_dfr(cs_results_ld, function(res) {
  if (is.null(res$agg)) return(NULL)
  tibble(
    partition  = res$partition,
    n_cohorts  = res$n_cohorts,
    n_treated  = res$n_treated,
    n_controls = res$n_controls,
    att        = res$agg$overall.att,
    se         = res$agg$overall.se,
    ci_lo      = res$agg$overall.att - 1.96 * res$agg$overall.se,
    ci_hi      = res$agg$overall.att + 1.96 * res$agg$overall.se,
    p_val      = 2 * pnorm(-abs(res$agg$overall.att / res$agg$overall.se))
  )
}) %>%
  mutate(
    sig       = case_when(
      p_val < 0.01 ~ "***",
      p_val < 0.05 ~ "**",
      p_val < 0.10 ~ "*",
      TRUE         ~ ""
    ),
    partition = factor(partition, levels = c("10-20", "20+"))
  ) %>%
  arrange(partition)
 

att_table_ld %>%
  transmute(
    `PGA Partition`  = partition,
    `Cohorts`        = n_cohorts,
    `Treated Munis`  = n_treated,
    `Control Munis`  = n_controls,
    `ATT`            = round(att,  4),
    `Std. Error`     = round(se,   4),
    `95% CI`         = paste0("[", round(ci_lo, 4), ", ", round(ci_hi, 4), "]"),
    `p-value`        = round(p_val, 3),
    ` `              = sig
  ) %>%
  knitr::kable(
    booktabs = TRUE,
    align    = c("l","c","c","c","r","r","c","r","l"),
    caption  = "CS ATT by PGA Partition | low-dose (4-10) as controls | doubly robust"
  )
```

As a robustness check, we re-estimate the model replacing the never-treated municipalities with those in the lowest PGA partition (4-10 %g) as the control group. The results, are consistnet. Both ATT estimates remain statistically significant and retain their respective signs: the moderate-intensity partition (10-20 %g) yields a positive and precisely estimated effect of 0.496 log points, while the high-intensity partition (above 20 %g) continues to exhibit a negative effect of −0.506 log points, significant at the 5% level. The standard errors are virtually unchanged relative to the baseline specification, and the confidence intervals overlap closely with those reported in the table above.

This stability is reassuring on two fronts. First, it suggests that the never-treated control group was not driving the results through some idiosyncratic feature of municipalities that experienced no seismic activity during the sample period. Second, it lends further credibility to the identifying assumption: municipalities exposed to low-intensity shocks appear to provide a valid counterfactual for more severely affected units, implying that the parallel trends assumption extends across treatment intensities and is not an artifact of the choice of control group which we can observe in the following plots.

```{r}
#| fig.width: 6
#| fig.height: 4
#| fig.align: "center"
#| out.width: "90%"

es_plots_ld <- map(cs_results_ld, function(res) {
  if (is.null(res$es)) return(NULL)
 
  ggdid(res$es) +
    labs(
      title    = paste("Event Study — PGA Partition:", res$partition),
      subtitle = paste0("CS (2021) | low-dose (4-10) as controls | ",
                        res$n_treated, " treated, ",
                        res$n_controls, " control munis"),
      x        = "Quarters relative to earthquake",
      y        = "ATT (Inverse hyperbolic sine remittances)"
    ) +
    theme_bw() +
    theme(plot.title    = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))
}) %>%
  compact()
 
walk(es_plots_ld, print)
```

The event-study plots for this alternative specification tell a consistent story. The post-treatment dynamics mirror those of the baseline model closely, with the positive trajectory for the moderate-intensity partition and the negative response for the high-intensity partition both preserved. Notably, the pre-treatment fit improves relative to the baseline specification: the minor deviations from parallel trends observed in the lowest PGA partition under the never-treated control group are no longer present, and the pre-period coefficients sit closer to zero across both partitions. This is consistent with the intuition that municipalities exposed to low-intensity shocks follow a more similar pre-treatment remittance trend to those exposed to moderate.

```{r}
ggplot(att_table_ld, aes(x = partition, y = att)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi),
                width = 0.15, color = "steelblue", linewidth = 0.8) +
  geom_point(color = "steelblue", size = 3) +
  geom_text(aes(label = sig, y = ci_hi + 0.01),
            size = 5, color = "steelblue") +
  labs(
    title    = "CS ATT by PGA Partition (low-dose controls)",
    subtitle = "Callaway & Sant'Anna (2021) | low-dose (4-10) as controls | doubly robust",
    x        = "PGA Partition (%g)",
    y        = "ATT (Inverse hyperbolic sine remittances)"
  ) +
  theme_bw() +
  theme(plot.title    = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```


## Testing for Spillover Effects {#sec-spillover}

### Geographic Distances

To test whether earthquake exposure in neighboring municipalities generates remittance spillovers through pure geographic proximity, we compute a geographic PGA exposure variable for each municipality using an inverse-distance-weighted spatial lag of the treatment dose vector:

$$\text{geo\_pga}_i = \sum_{j \neq i} w^{\text{geo}}_{ij} \cdot \mathbf{1}[\text{PGA}_j > \tau]$$

where $w^{\text{geo}}_{ij} = \frac{d_{ij}^{-1}}{\sum_{k \neq i} d_{ik}^{-1}}$ is the row-normalized inverse geodesic distance between municipal centroids, constructed from the geographic weight matrix $W^{\text{geo}}$.

```{r geographic pga}
# define our 'dose' matrix
dose_lookup <- panel %>%
  mutate(muni_id = paste(state, municipality, sep = "_")) %>%
  group_by(muni_id) %>%
  summarise(dose = max(mean_pga, na.rm = TRUE))

# call our geographic matrix
W_geo <- readRDS("data/W_geo.rds")

# match dimensions to do matrix multiplication
dose_vector_geo <- setNames(rep(0, nrow(W_geo)), 
                            rownames(W_geo))
matched_geo <- intersect(rownames(W_geo), # find shared muni names
                         dose_lookup$muni_id)
dose_vector_geo[matched_geo] <- dose_lookup$dose[match(matched_geo,
                                                       dose_lookup$muni_id)]
# fill in PGA values, missing assigned 0

# sanity check
dim(W_geo)[2] == length(dose_vector_geo) # TRUE

# geography matrix * dose_vector
geo_pga = as.vector(W_geo %*% dose_vector_geo)

# store in tibble df
geo_pga_df <- tibble(
  muni_id = rownames(W_geo),
  geo_pga = geo_pga
)
```

We then estimate:

$$\log(\text{remit}_{it}) = \alpha_i + \gamma_t + \beta \cdot \text{post}_{it} \times \widetilde{\text{geo\_pga}}_i + \varepsilon_{it}$$

where $\widetilde{\text{geo\_pga}}_i$ is mean-centered, and standard errors are clustered at the municipality level.

```{r geographic spillover estimate}
spillover_df <- panel_cs %>%
  left_join(geo_pga_df, by = "muni_id") %>%
  filter(G_cs == 0, !is.na(geo_pga)) %>%
  mutate(
    post      = as.integer(time_index >= min(muni_meta$G)),
    geo_pga_c = geo_pga - mean(geo_pga, na.rm = TRUE)
  )

spillover_mod_geo <- feols(
  asinhremit ~ post:geo_pga_c | muni_num + time_index,
  data    = spillover_df,
  cluster = ~muni_num
)

summary(spillover_mod_geo)
```

The univariate geographic specification yields $\hat{\beta} = 0.097$ ($p < 0.001^{***}$), indicating that municipalities geographically proximate to severely shaken areas receive significantly higher remittance inflows in the post-earthquake period. Taken in isolation, this result is consistent with a pure proximity-based spillover story, which perhaps reflects shared regional labor-market disruption or aggregate demand shocks that span administrative boundaries.

### Migration Network

Geographic proximity alone cannot distinguish between two competing mechanisms: 

  (i) direct geographic contagion, and 
  (ii) diaspora network propagation, 

where migrant communities in the United States redirect remittances toward origin municipalities that are socially connected to affected areas, regardless of physical distance. To isolate the network channel, we compute a migration-network PGA exposure variable:

$$\text{network\_pga}_i = \sum_{j \neq i} w^{\text{net}}_{ij} \cdot \mathbf{1}[\text{PGA}_j > \tau]$$

where $W^{\text{net}} = MM^\top$ is constructed from the municipality $\times$ US-state migration share matrix $M$, row-normalized so that $w^{\text{net}}_{ij}$ reflects the share of common migrant-destination overlap between municipalities $i$ and $j$.

```{r network pga}
# call in migration network 
W_network <- readRDS("data/W_network.rds")

# same dose vector as geographic section: max PGA per municipality
dose_vector_net <- setNames(rep(0, nrow(W_network)), rownames(W_network))
matched_net     <- intersect(rownames(W_network), dose_lookup$muni_id)
dose_vector_net[matched_net] <- dose_lookup$dose[match(matched_net,
                                                        dose_lookup$muni_id)]

# sanity check
dim(W_network)[2] == length(dose_vector_net) # TRUE

# network matrix * dose vector
network_pga_df <- tibble(
  muni_id     = rownames(W_network),
  network_pga = as.vector(W_network %*% dose_vector_net)
)
```

```{r network spillover estimate}
# join on CS panel, create 'post' and mean 0 pga
spillover_df_net <- panel_cs %>%
  left_join(network_pga_df, by = "muni_id") %>%
  filter(G_cs == 0, !is.na(network_pga)) %>%
  mutate(
    post          = as.integer(time_index >= min(muni_meta$G)),
    network_pga_c = network_pga - mean(network_pga, na.rm = TRUE)
  )

# TWFE reg
spillover_mod_net <- feols(
  asinhremit ~ post:network_pga_c | muni_num + time_index,
  data    = spillover_df_net,
  cluster = ~muni_num
)

summary(spillover_mod_net)
```

The univariate network specification yields $\hat{\beta} = 0.162$ ($p = 0.0397$, $^{*}$), confirming that municipalities with stronger diaspora network linkages to earthquake-affected areas experience increased remittance inflows post-event. This is consistent with the altruistic remittance hypothesis, wherein US-based migrants respond to earthquake news in socially connected origin communities, not solely in the geographically nearest ones.

### Joint Test

The univariate results are both individually significant, but geographic proximity and migration network connectivity are, by construction, correlated. Municipalities that share borders tend also to share migration corridors. To assess which mechanism drives the spillover, we estimate both simultaneously:

$$\log(\text{remit}_{it}) = \alpha_i + \gamma_t + \beta_1 \left( \text{post}_{it} \cdot \widetilde{\text{network\_pga}}_i \right) + \beta_2 \left( \text{post}_{it} \cdot \widetilde{\text{geo\_pga}}_i \right) + \varepsilon_{it}$$

```{r spillover test}
spillover_df_joint <- panel_cs %>%
  left_join(network_pga_df, by = "muni_id") %>%
  left_join(geo_pga_df,     by = "muni_id") %>%
  filter(G_cs == 0, !is.na(network_pga), !is.na(geo_pga)) %>%
  mutate(
    post          = as.integer(time_index >= min(muni_meta$G)),
    network_pga_c = network_pga - mean(network_pga, na.rm = TRUE),
    geo_pga_c     = geo_pga     - mean(geo_pga,     na.rm = TRUE)
  )

spillover_mod_joint <- feols(
  asinhremit ~ post:network_pga_c + post:geo_pga_c | muni_num + time_index,
  data    = spillover_df_joint,
  cluster = ~muni_num
)

summary(spillover_mod_joint)
```

From the joint specification results, we notice that the network exposure coefficient remains stable at $\hat{\beta}_1 = 0.162$ ($p = 0.043^{*}$), while the geographic exposure coefficient collapses to $\hat{\beta}_2 = -0.005$ ($p = 0.882$), statistically indistinguishable from zero. This suggests that, although the geographic distances are much more significant univariately, once network exposure is controlled for the geography contributes no additional explanatory power. The spillover mechanism is therefore not driven by physical distance or regional economic contagion, but exclusively by the structure of the migrant diaspora network. 

Under SUTVA, control municipalities should be unaffected by treatment; the significant network spillover coefficient ($\hat{\beta}_1 = 0.162$, $p = 0.043$) indicates a partial violation structured along diaspora lines. 

This implies that the CS estimates may be understating how much remittances fall in the directly-hit municipalities, because some control municipalities are receiving network-driven remittance boosts that make the treated-control gap appear smaller than it actually may be.


### Visualization

```{r, echo=FALSE}
#| fig-cap: "Spillover Exposure: Network vs Geography"
#| label: fig-spillover-maps

# chiapos-only dose vector for map
dose_lookup_chiapas <- panel %>%
  mutate(muni_id = paste(state, municipality, sep = "_")) %>%
  filter(grepl("chiapas", event, ignore.case = TRUE)) %>%
  group_by(muni_id) %>%
  summarise(dose = max(mean_pga, na.rm = TRUE), .groups = "drop")

dose_vector_geo_chiapas <- setNames(rep(0, nrow(W_geo)), rownames(W_geo))
matched_geo_c <- intersect(rownames(W_geo), dose_lookup_chiapas$muni_id)
dose_vector_geo_chiapas[matched_geo_c] <- dose_lookup_chiapas$dose[
  match(matched_geo_c, dose_lookup_chiapas$muni_id)
]

geo_pga_chiapas_df <- tibble(
  muni_id = rownames(W_geo),
  geo_pga_chiapas = as.vector(W_geo %*% dose_vector_geo_chiapas)
)

# admin borders
mex_muni <- st_read("data/earthquakes/mex_admin2/mex_admin2.shp", quiet = TRUE) %>%
  st_transform("EPSG:4326") %>%
  mutate(adm1_name = case_when(
    adm1_name == "Coahuila de Zaragoza"            ~ "Coahuila",
    adm1_name == "Distrito Federal"                ~ "Ciudad de México",
    adm1_name == "Veracruz de Ignacio de la Llave" ~ "Veracruz",
    adm1_name == "Michoacán de Ocampo"             ~ "Michoacán",
    adm1_name == "Querétaro de Arteaga"            ~ "Querétaro",
    TRUE                                            ~ adm1_name
  ))

# crop to Mexico 
mex_admin0 <- st_read("data/earthquakes/mex_admin0/mex_admin0.shp", 
                      quiet = TRUE) %>% 
  st_transform("EPSG:4326")

# state borders
mex_states <- st_read("data/earthquakes/mex_admin1/mex_admin1.shp", 
                      quiet = TRUE) %>%
  st_transform("EPSG:4326")

# join both exposure variables to municipality geometries
map_spillover <- mex_muni %>%
  mutate(muni_id = paste(adm1_name, adm2_name, sep = "_")) %>%
  left_join(network_pga_df,      by = "muni_id") %>%
  left_join(geo_pga_chiapas_df,  by = "muni_id")

# network exposure plot
p_net_exp <- ggplot() +
  geom_sf(data = map_spillover, aes(fill = network_pga), color = NA) +
  geom_sf(data = mex_states, fill = NA, color = "grey30", linewidth = 0.3) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1,
                       name = "Network\nexposure", na.value = "grey88") +
  coord_sf(xlim = c(-119, -86), ylim = c(14, 33)) +
  labs(title = "Migration Network", x = NULL, y = NULL) +
  theme_bw() +
  theme(plot.title  = element_text(hjust = 0.5),
        axis.text   = element_blank(),
        axis.ticks  = element_blank())

# geographic exposure plot
p_geo_exp <- ggplot() +
  geom_sf(data = map_spillover, aes(fill = geo_pga_chiapas), color = NA) +
  geom_sf(data = mex_states, fill = NA, color = "grey30", linewidth = 0.3) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1,
                       name = "Geographic\nexposure", na.value = "grey88") +
  coord_sf(xlim = c(-119, -86), ylim = c(14, 33)) +
  labs(title = "Geographic Distance", x = NULL, y = NULL) +
  theme_bw() +
  theme(plot.title  = element_text(hjust = 0.5),
        axis.text   = element_blank(),
        axis.ticks  = element_blank())

# side-by-side
p_net_exp + p_geo_exp +
  plot_annotation(
    title    = "Spillover Exposure: Network vs Geography",
    subtitle = "2017 Chiapas M8.2 | never-treated municipalities",
    theme    = theme(plot.title    = element_text(hjust = 0.5),
                     plot.subtitle = element_text(hjust = 0.5))
  )
```

Inspecting Figure @fig-spillover-maps visually, the geographic exposure graph shows a smooth decay pattern centered on the Chiapas epicenter, while the network exposure graph disperses across traditional high-emigration states in central and western Mexico, far from the earthquake zone. The two graphs highlight largely different municipalities, confirming that network and geographic exposure are empirically separable and that the joint regression is identifying distinct channels.

# Plots: Remittances and Earthquakes

## 2017 M8.2 Chiapas

@fig-earthquake-raster shows the PGA spread of the 2017 Chiapas M8.2 earthquake, with shaking concentrated along the southern coast and decaying towards the north. Municipality borders are colored by PGA partition, defining the four treatment intensity bins used in the Callaway-Sant'Anna estimation.

@fig-earthquake-remittances plots the percentage change in municipal remittance inflows relative to the pre-earthquake quarter (Q2 2017) for the four quarters immediately following the event. Two patterns are visible: remittances increase sharply in several municipalities in Q+0 and Q+1, consistent with an immediate diaspora response, but the picture becomes more heterogeneous by Q+2 and Q+3 as both large gains and losses emerge across Mexico. Notably, the response is not spatially concentrated around the epicenter, as municipalities far from the earthquake zone also show significant changes, which motivates the spillover analysis in Section @sec-spillover.

### Earthquake Raster

```{r, echo=FALSE}
#| fig-cap: "PGA %g of M8.2 Earthquake near Tehuantepec, Mexico in 2017"
#| label: fig-earthquake-raster

library(terra)
library(tidyterra)
library(sf)

# crop to Mexico 
mex_admin0 <- st_read("data/earthquakes/mex_admin0/mex_admin0.shp", 
                      quiet = TRUE) %>% 
  st_transform("EPSG:4326")

# state borders
mex_states <- st_read("data/earthquakes/mex_admin1/mex_admin1.shp", 
                      quiet = TRUE) %>%
  st_transform("EPSG:4326")

# PGA raster
pga_rast <- rast("output/2017_M8.2_chiapas/shakemap_pga_mexico_masked.tif") %>%
  project("EPSG:4326")

# Epicenter
epicenter_2017 <- data.frame(lon = -93.899, 
                             lat = 15.022)

# Municipality shapefile (with state name cleaning)
mex_muni <- st_read("data/earthquakes/mex_admin2/mex_admin2.shp", quiet = TRUE) %>%
  st_transform("EPSG:4326") %>%
  mutate(adm1_name = case_when(
    adm1_name == "Coahuila de Zaragoza"            ~ "Coahuila",
    adm1_name == "Distrito Federal"                ~ "Ciudad de México",
    adm1_name == "Veracruz de Ignacio de la Llave" ~ "Veracruz",
    adm1_name == "Michoacán de Ocampo"             ~ "Michoacán",
    adm1_name == "Querétaro de Arteaga"            ~ "Querétaro",
    TRUE                                            ~ adm1_name
  ))

# PGA partition borders (treated municipalities only, colored by partition)
panel <- read_csv("data/panel_remittances_pga.csv")

map_borders <- mex_muni %>%
  left_join(
    panel %>%
      filter(grepl("chiapas", event, ignore.case = TRUE), mean_pga > 4) %>%
      mutate(pga_bin = case_when(
        mean_pga <= 10 ~ "4-10",
        mean_pga <= 20 ~ "10-20",
        mean_pga <= 35 ~ "20-35",
        TRUE           ~ "35+"
      )) %>%
      select(state, municipality, pga_bin),
    by = c("adm1_name" = "state", "adm2_name" = "municipality")
  ) %>%
  filter(!is.na(pga_bin)) %>%
  mutate(pga_bin = factor(pga_bin, levels = c("4-10", "10-20", "20-35", "35+")))


p_pga <- ggplot() +
  # base layer: Mexico vector object
  geom_sf(data = mex_muni, fill = "grey92", color = NA) +
  # PGA raster on top
  geom_spatraster(data = pga_rast) +
  # state borders 
  geom_sf(data = mex_states, fill = NA, color = "grey30", linewidth = 0.35) +
  # PGA partition municipality borders
  geom_sf(data = map_borders, aes(color = pga_bin),
          fill = NA, linewidth = 0.7) +
  geom_point(data = epicenter_2017, aes(x = lon, y = lat),
             shape = 24, fill = "yellow", color = "black", size = 3) +
  scale_fill_distiller(
    palette   = "YlOrRd",
    direction = 1,
    name      = "PGA (%g)",
    na.value  = "transparent"
  ) +
  scale_color_manual(
    values = c("4-10"  = "#FEE08B", "10-20" = "#FDAE61",
               "20-35" = "#F46D43", "35+"   = "#D73027"),
    guide  = "none"
  ) +
  coord_sf(xlim = c(-119, -86), ylim = c(14, 33)) +
  labs(title = "2017 Chiapas M8.2 (PGA %g)", x = NULL, y = NULL) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

p_pga
```

### Remittances Across Municipalities

```{r, echo=FALSE}
#| fig-cap: "Percentage Change in Remittances, post 2017 Tehuantepec Earthquake"
#| label: fig-earthquake-remittances

pre <- panel %>%
  filter(year == 2017, quarter == 2) %>%
  select(state, municipality, remit_pre = remittances_musd)

post <- panel %>%
  filter(
    (year == 2017 & quarter %in% c(3, 4)) |
    (year == 2018 & quarter %in% c(1, 2))
  ) %>%
  mutate(label = case_when(
    year == 2017 & quarter == 3 ~ "Q+0  (Q3 2017)",
    year == 2017 & quarter == 4 ~ "Q+1  (Q4 2017)",
    year == 2018 & quarter == 1 ~ "Q+2  (Q1 2018)",
    year == 2018 & quarter == 2 ~ "Q+3  (Q2 2018)"
  ))

remit_change <- post %>%
  left_join(pre, by = c("state", "municipality")) %>%
  mutate(pct_change = pmax(pmin(
    (remittances_musd - remit_pre) / (remit_pre + 1) * 100,
    60), -60))

map_remit <- mex_muni %>%
  left_join(remit_change,
            by = c("adm1_name" = "state", "adm2_name" = "municipality"))

# Add PGA bin to map_remit so we can draw partition borders
map_remit_borders <- map_remit %>%
  left_join(
    panel %>%
      filter(year == 2017, quarter == 3) %>%
      mutate(pga_bin = case_when(
        mean_pga >  4 & mean_pga <= 10 ~ "4-10",
        mean_pga > 10 & mean_pga <= 20 ~ "10-20",
        mean_pga > 20 & mean_pga <= 35 ~ "20-35",
        mean_pga > 35                  ~ "35+",
        TRUE                           ~ NA_character_
      )) %>%
      select(state, municipality, pga_bin),
    by = c("adm1_name" = "state", "adm2_name" = "municipality")
  )

p_remit <- ggplot() +
  # All municipalities: fill = % remittance change
  geom_sf(data  = map_remit_borders %>% filter(!is.na(label)),
          aes(fill = pct_change),
          color = NA, linewidth = 0.05) +
  geom_sf(data  = mex_states,
          fill  = NA, color = "grey20", linewidth = 0.35) +
  scale_fill_gradient2(
    low      = "#2166AC",
    mid      = "white",
    high     = "#D73027",
    midpoint = 0,
    limits   = c(-60, 60),
    oob      = scales::squish,
    name     = "% change\nvs Q2 2017",
    na.value = "grey88"
  ) +
  facet_wrap(~ label, ncol = 2) +
  coord_sf(xlim = c(-119, -86), ylim = c(14, 33)) +
  labs(
    title    = "Remittance % Change After 2017 Chiapas M8.2",
    x = NULL, y = NULL
  ) +
  theme_bw() +
  theme(
    plot.title       = element_text(hjust = 0.5),
    plot.subtitle    = element_text(hjust = 0.5),
    strip.background = element_rect(fill = "grey92"),
    legend.position  = "right",
    axis.text  = element_blank(),
    axis.ticks = element_blank()

  ) 

p_remit
```
