---
title: "Geospatial Project"
author: "Joan Costa Rodríguez, Francesco Zucca, Xènia Alemany Rovira, Michael Fehl"
date: today
format:
  pdf:
    toc: true
    number-sections: true
    geometry: margin=2.5cm
    fontsize: 12pt
bibliography: "references.bib"
execute:
  echo: true
  warning: false
---

```{r libraries, results='hide'}
rm(list=ls())
library(readxl)        # read excel files
library(stringr)       # string operations
library(terra)         # raster operations
library(exactextractr) # exact_extract() for distances
library(gdistance)     # fast distance() formula for geo data
library(sf)            # vector operations
library(spData)        # geo data
library(tidyverse)     # NOTE: Conflict -- tidyr::extract() masks terra::extract()
library(spDataLarge)   # large geo data
library(ggrepel)       # add labels to ggplot
library(tidyterra)     # visualize geospatial using tidy syntax
library(patchwork)     # patch tg ggplot objects
```

# Abstract

Does the severity of earthquakes determine how much additional remittance income a municipality receives after the disaster happened? Unlike existing macro-level studies that rely on state data, this paper exploits more granular municipal level data to evaluate the impact of earthquake severity on post-disaster remittances. At this local level spatial spillovers between neighboring municipalities may violate SUTVA, introducing attenuation bias and artificially dampening the observed regional differences.

To investigate this, we construct a municipality-quarter panel of Mexican remittance inflows from Banco de México (2013 - 2025) [@banxico2025], merged with Peak Ground Acceleration (PGA) data from the USGS Earthquake Hazard Program [@usgs2026]. Using this dataset, we estimate a Two-Way Fixed Effects (TWFE) Difference-in-Differences model - **actually we probably need to use Callaway Sant' Anna :(** -, where treatment intensity is defined as the average PGA experienced by each municipality.

# Introduction

In 2024 alone, Mexico received more that 67 billion current US dollars in remittances, making it the second country in the world by nominal remittance inflows, behind India [@worldbank2026]. When a shock hits the family in the municipality of origin, remittances act as an insurance mechanism, increasing flows for the following months [@lucas_stark1985, @bettin_jallow_zazzaro2023]. Mexico is one of the most seismically active countries in the world because of its location over three of the major tectonic plates [@rhea2011]. Particularly, on 7 September 2017, a magnitude 8.2 earthquake struck the coast of Chiapas, followed after around two weeks by a magnitude 7.1 event near Puebla, the latter one causing the collapse of 44 buildings and severely damaging another 600 [@singh2018].

[@mohapatra2012] use cross-country data and find that remittances increase after natural disasters, especially where there is a large stock of migrant population. [@bettin2025] use monthly bilateral remittance data from Italy to 81 countries and estimate an immediate 2% increase following these type of events, reverting back to the normal flow after around one year. While precise, these papers aggregate at the country level and rely on a binary treatment definition. Our approach, in contrast, utilizes more granular data and a continuous measure of treatment. This allows us to test two competing outcomes: either we will identify a stronger, localized response effect in the hardest-hit municipalities, or we will find that network effects completely absorb these granular differences.

# Data

## Remittance Data

The outcome of our analysis is quarterly remittance inflows from Banco de México SIE table CE166 - *Ingresos por Remesas, distribución por municipio* -, covering all Mexican municipalities from Q1 2013 to Q4 2025.

```{r remittance data}
# Read the excel skipping the metadata rows
remit_raw <- read_excel(
  "data/remittances/remittances.xlsx",
  skip = 8,
  col_types = "text"
)
```

We convert the dataset to long format and convert the date column in date format, adding columns for years and quarters to improve readability.

```{r remittance data long}
# Convert to long format
remit_long <- remit_raw %>%
  select(-`Information type`) %>% # Drop levels columns
  pivot_longer(
    cols      = -c(Title, Serie),
    names_to  = "quarter_label",
    values_to = "remittances_musd"
  ) %>%
  mutate(remittances_musd = as.numeric(remittances_musd))

# Convert quarter column to date format
remit_long <- remit_long %>%
  mutate(
    date    = as.Date(as.numeric(quarter_label), origin = "1899-12-30"),
    year    = as.integer(format(date, "%Y")),
    quarter = ceiling(as.integer(format(date, "%m")) / 3)
  ) %>%
  select(-quarter_label) # drop the old quarter column
```

Being interested only in municipality level data, we drop the State aggregate rows. Moreover, to improve readability once more, we add the corresponding state column to the left of the municipality one.

```{r remittance data cleaning}
# Tidy up municipality names
remit_long <- remit_long %>%
  # keep only remittances rows (drop state-level aggregates)
  filter(str_detect(Title, "Distribution by municipality")) %>%
  mutate(
    state        = str_trim(str_extract(Title, "(?<=, )[^,]+(?=, [^,]+$)")),
    municipality = str_trim(str_extract(Title, "[^,]+$"))
  ) %>%
  select(-Title)  # drop the original Title column

# Drop unused columns and rearrange columns
remit_long <- remit_long %>%
  select(-date, -Serie) %>%
  select(state, municipality, year, quarter, remittances_musd)
```


## Earthquake Data

```{r}


```


# Spillover Effects

## Migration Networks

To get a measure of potential spillover effects, we can use a manually constructed weighting matrix of mexican migration from a given Mexican municipality to each US state. We can then multiply this matrix by its transpose to get a municipality x municipality matrix to try and understand the connectedness of certain Mexican municipalities based on their shares US migration destinations. Theoretically, this means that certain Mexican municipalities that are not close in geographic distance can still propagate shocks through these migration networks.

```{r import weighting matrix}
M <- read_excel("C:/Users/T14/1BSE/8Thesis/1_network_estimation/2_migration_matrix_estimation/migration_weighting_matrices_2/AVG_WEIGHTING_MATRIX.xlsx") 
muni_ids <- M[[2]]   # store IDs
state_ids <- M[[1]]  
muni_labels <- paste(state_ids, muni_ids, sep = "_") # generate unique label for each muni

M <- as.matrix(M[c(-1, -2)]) # drop names, convert to matrix
dim(M) # (2638 x 51) <==> (MX muni x US states) (District of Columbia gives us +1)

# transpose matrix
M_prime <- t(M)

# MM' gives our (muni x muni) network
W_network <- M %*% M_prime
dim(W_network)

# place muni names back in
rownames(W_network) <- muni_labels
colnames(W_network) <- muni_labels

# municipality is not its own neighbor
diag(W_network) <- 0

# row normalize: add to one
W_network <- W_network/rowSums(W_network)

# sanity check
as.data.frame(W_network) %>% 
  slice(1:4) %>% 
  select(1:2)

# largest values in a row = strongest network neighbors
# sort(W_network["Michoacan_Coeneo",], decreasing = TRUE)[1:5]
```

:::{.callout-note}
Note that we row-normalize our final matrix here; what is important is that this final product is not a symmetric matrix. The interpretation is that for a given row $i$ (municipality), we get the probability distribution across all other municipalities ($\forall_{i \neq j}$. We must include it in this form in order to properly use it as a variable in our regression later on.
:::

```{r clean up environment, results='hide', echo=FALSE}
rm(M, M_prime, muni_ids, state_ids)
```

This spillover matrix serves to approximately measure the "connectedness" between every pair of municipalities in Mexico, based on their historical migration patterns to the US. The idea is that if an earthquake hits Asientos in Aguascalientes, the migration community in (say) Texas that comes from this municipality will send more money home. But if that same community in Texas also has a large network of migrants from Coeneo in Michoacan, they might discuss the earthquake; consequently, migrants from Coeneo may send additional remittances back home as feel a heightened sense of their own family's vulnerability; in other words, they anticipate risk associated with disaster in another municipality. 


## Geographic Distance

Furthermore, we may suspect that neighboring municipalities in Mexico would contribute to this spillover effect. Theoretically, if an earthquake hits a municipality hard enough they might rely on aid from neighboring towns as local residents may be out of work and cannot afford to survive after such a disaster. Thus, we can create another matrix, this time using a simple weighting matrix based on geographic distance to the affected municipalit(ies). 

To do so, we can simply use our previous weighting matrix and intialize the matrix by assigning zeros to all (municipality x municipality) pairs. 


```{r geographic matrix}
empty_network = W_network * 0 # init empty matrix

```



# Regression: Callaway Sant-Anna

parallel trends ass

what regressors we include in the model


